Project 1 for Complex Adaptive Systems

### 2018-02-20 paper.tex
final section

@@

\begin{figure*}[!t]
\centering
\subfloat[]{\includegraphics[width=3.5in]{Fig4a}%
\label{te} }

\hfill
\subfloat[]{\includegraphics[width=3.5in]{Fig4b}%
\label{fig_second_case} }
\caption{(a) Top-down, $T_{M\rightarrow X}$ (solid), and bottom-up, $T_{X\rightarrow M}$ (dashed), causal information transfer as a function of the global coupling parameter $\epsilon$ for coupled logistic maps. (b) The difference (Top-down minus bottom-up) in causal information transfer for $30$ coupled systems as well as the mean as a function of $\epsilon$ (dashed). Vertical line at $\epsilon = 0.3$ illustrates the directional reversal of causal information transfer.}
\label{te_avg}
\end{figure*}

\subsection{Direction of Information Flow}
\noindent In order to assess that direction of information flow, from Global to Local scales and vice versa, the authors in \cite{Walker} take advantage of {\it Transfer Entropy}. Transfer Entropy is defined mathematically in \cite{Schreiber} as
$$T^{(k)}_{Y \rightarrow X} = \sum_n p(x_{n+1}, x_n^{(k)}, y_n^{(k)}) \log \left[ \frac{p(x_{n+1}|x_n^{(k)}, y_n^{(k)})}{p(x_{n+1}|x_n{(k)}}\right]$$
Where $x_n^{(k)}$ is an embedded state in the $k$-dimensional phase spate defined as $x_n^{(k)} = (x_n, \cdots, x_{n-k+1})$. \\

\noindent Although Transfer Entropy is a fine candidate to estimate directionality of this coupling, recent work has focused on improving estimation methods. In their 2015 paper, Gencaga et. al propose new "statistically robust" methodology to estimate Transfer Entropy \cite{recipe}. Relevant here, is the considerable improvement for continuous processes, where discretizing the data can lead to a distortion in the underlying probability distribution. For this reason, we choose to use the state of the art TransferEntropy package in R and claim based on this work that these estimates of Transfer Entropy are to be preferred. \\

@@

\subsection{Additional Analysis}
Here we continue our exploration of the relationship between local and global population dynamics.

>>>>

\begin{figure*}[!t]
\centering
\subfloat[]{\includegraphics[width=3.5in]{Fig4a}%
\label{te}}
\hfill
\subfloat[]{\includegraphics[width=3.5in]{Fig4b}%
\label{fig_second_case} }
\caption{(a) Top-down, $T_{M\rightarrow X}$ (solid), and bottom-up, $T_{X\rightarrow M}$ (dashed), causal information transfer as a function of the global coupling parameter $\epsilon$ for coupled logistic maps. (b) The difference (Top-down minus bottom-up) in causal information transfer for $30$ coupled systems as well as the mean as a function of $\epsilon$ (dashed). Vertical line at $\epsilon = 0.3$ illustrates the directional reversal of causal information transfer.}
\label{te_avg}
\end{figure*}

@@

\subsection{Direction of Information Flow}
\noindent In order to assess that direction of information flow, from Global to Local scales and vice versa, the authors in \cite{Walker} take advantage of {\it Transfer Entropy}. Transfer Entropy is defined mathematically in \cite{Schreiber} as
$$T^{(k)}_{Y \rightarrow X} = \sum_n p(x_{n+1}, x_n^{(k)}, y_n^{(k)}) \log \left[ \frac{p(x_{n+1}|x_n^{(k)}, y_n^{(k)})}{p(x_{n+1}|x_n{(k)}}\right]$$
Where $x_n^{(k)}$ is an embedded state in the $k$-dimensional phase spate defined as $x_n^{(k)} = (x_n, \cdots, x_{n-k+1})$. \\
\begin{figure}[t]
\centering
 \includegraphics[width=90mm]{Fig5}

  \caption{Population trajectories for an individual (blue) and the mean (red) over 100 time steps. Dynamics are shown for $\epsilon = 0.075$ ($T_{X\rightarrow M} > T_{M\rightarrow X}$), $\epsilon = 0.3$ ($T_{X\rightarrow M} = 0 = T_{M\rightarrow X}$) and $\epsilon = 0.4$ ($T_{X\rightarrow M} < T_{M\rightarrow X}$. }
 \label{timeseries}
\end{figure}

\noindent Although Transfer Entropy is a fine candidate to estimate directionality of this coupling, recent work has focused on improving estimation methods. In their 2015 paper, Gencaga et. al propose new "statistically robust" methodology to estimate Transfer Entropy \cite{recipe}. Relevant here, is the considerable improvement for continuous processes, where discretizing the data can lead to a distortion in the underlying probability distribution. For this reason, we choose to use the state of the art TransferEntropy package in R and claim based on this work that these estimates of Transfer Entropy are to be preferred. \\
x

@@

\subsection{Additional Analysis}
\noindent Here we continue our exploration of the relationship between local and global population dynamics. We will use the metric displayed in Figure \ref{te_avg} $T_{M\rightarrow X} - T_{X\rightarrow M}$ to discuss the direction of causal inforamtion transfer between the two scales. We break our discussion into 3 cases.
\begin{itemize}
\item[i)] For small values of the global coupling strength, say $\epsilon < 0.3$, Figure \ref{te_avg} shows that our metric takes negative values. This indicates that causal information is flowing from the local to the global scale. The top row of Figure \ref{timeseries} shows that the individual population is chaotic, but the mean trajectory oscilates roughly between 3 values ($0.4$, $0.6$ and $0.6$). This is what gives the red points their behavior in Figure \ref{prettypicture}.
\item[ii)] For $\epsilon \approx 0.3$, the system synchronizes. In fact, we can compute the correlation between the trajectories in the middle row of Figure \ref{timeseries}, and see that it is nearly equal to one (0.9879)$. Based on Figure \ref{prettypicture} we sould expect these series to oscillate between precisely two values. This particular run had four attractors (black stars) instead of two.
\item[iii)] For large values of the global coupling strength, the global dynamics begin to dominate. There is still a near perfect correlation between $M$ and $X$, but now even the global dynamics (the mean) behaves chaotically.
\end{itemize}
For small values of the Global Coupling Strength, say $\epsilon < 0.3$, this metric takes negative values, indicating that information is flowing from the local to the global scale. The top row of Figure \ref{timeseires} shows that the individual populations appear chaotic, but the mean trajectory appears periodic. This is similar to the behavior illus
